Concurrent Programming with CXL

### 2020-06-25 book/paper.tex
More index terms

There is a second version of the \texttt{synch} module that uses suspension
instead of spinlocks.
Figure~\ref{fig:suspension} shows the same \texttt{Lock} interface implemented
using suspension.
The interface is implemented as follows:
\begin{itemize}
\item A \texttt{Lock} maintains both a boolean indicating whether the
lock is taken and a list of contexts of processes that want to acquire the lock.
\item
\texttt{lock} sets the lock if available and suspends itself if not.
Note that \texttt{stop} is called within an \texttt{atomic} statement---this is
the only exception to an atomic statement running to completion.  While the
process is running no other processes can run, but when it suspends itself
it allows other processes to run.
\item
\texttt{unlock} checks to see if any processes are waiting to get the lock.
If so, it uses the \texttt{head} and \texttt{tail}
methods from the \texttt{list} module (see Appendix~\ref{ap:module})
to resume the first process that got
suspended and to remove its context from the list.
\end{itemize}
Selecting the first process is a design choice.  Another implementation could
have picked the last one, and yet another implementation could have used
\texttt{choose} to pick an arbitrary one.  Selecting the first is a common
choice in lock implementations as it prevents \emph{starvation}:
\index{starvation}
every process
gets a chance to obtain the lock (assuming every process eventually releases
it).  Chapter~\ref{ch:starvation} will talk more about starvation and how
to prevent it.

@@

\section*{Exercises}
\begin{problems}
\item Add read-only methods to the data structure in Figures~\ref{fig:lst1}
and~\ref{fig:lst2} that report the size of the list, the minimum value in the
list, the maximum value in the list, and the sum of the values in the list.
Are they linearizable?  If so, what are their linearization points?
\item A \emph{seqlock} consists of a lock and a version number.
An update operation acquires the lock, increments the version number, makes the
changes to the data structure, and then releases the lock.  A read-only operation
does not use the lock.  Instead, it retrieves the version number,
reads the data structure, and then checks if the
version number has not changed.  If it has, read-only operation is retried.
Use a seqlock to implement a bank much like Exercise~\ref{ex:bank}, with
one seqlock for the entire bank (i.e., no locks on individual accounts).
Method \texttt{transfer} is an update operation; method \texttt{total} is a
read-only operation.  Explain how a \texttt{seqlock} can lead to starvation.
\end{problems}

>>>>

There is a second version of the \texttt{synch} module that uses suspension
instead of spinlocks.
Figure~\ref{fig:suspension} shows the same \texttt{Lock} interface implemented
using suspension.
The interface is implemented as follows:
\begin{itemize}
\item A \texttt{Lock} maintains both a boolean indicating whether the
lock is taken and a list of contexts of processes that want to acquire the lock.
\item
\texttt{lock}
\index{lock}
sets the lock if available and suspends itself if not.
Note that \texttt{stop} is called within an \texttt{atomic} statement---this is
the only exception to an atomic statement running to completion.  While the
process is running no other processes can run, but when it suspends itself
it allows other processes to run.
\item
\texttt{unlock}
\index{unlock}
checks to see if any processes are waiting to get the lock.
If so, it uses the \texttt{head} and \texttt{tail}
methods from the \texttt{list} module (see Appendix~\ref{ap:module})
to resume the first process that got
suspended and to remove its context from the list.
\end{itemize}
Selecting the first process is a design choice.  Another implementation could
have picked the last one, and yet another implementation could have used
\texttt{choose} to pick an arbitrary one.  Selecting the first is a common
choice in lock implementations as it prevents \emph{starvation}:
\index{starvation}
every process
gets a chance to obtain the lock (assuming every process eventually releases
it).  Chapter~\ref{ch:starvation} will talk more about starvation and how
to prevent it.

@@

\section*{Exercises}
\begin{problems}
\item Add read-only methods to the data structure in Figures~\ref{fig:lst1}
and~\ref{fig:lst2} that report the size of the list, the minimum value in the
list, the maximum value in the list, and the sum of the values in the list.
Are they linearizable?  If so, what are their linearization points?
\item A \emph{seqlock}
\index{seqlock}
consists of a lock and a version number.
An update operation acquires the lock, increments the version number, makes the
changes to the data structure, and then releases the lock.  A read-only operation
does not use the lock.  Instead, it retrieves the version number,
reads the data structure, and then checks if the
version number has not changed.  If it has, read-only operation is retried.
Use a seqlock to implement a bank much like Exercise~\ref{ex:bank}, with
one seqlock for the entire bank (i.e., no locks on individual accounts).
Method \texttt{transfer} is an update operation; method \texttt{total} is a
read-only operation.  Explain how a \texttt{seqlock} can lead to starvation.
\end{problems}
